---
title: "*In introduction to atomic vectors*"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    df_print: default
    includes:
      in_header: resizer.html
runtime: shiny_prerendered
description: >
  Learn about atomic vectors, the basic data types in R.
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(skimr)
library(knitr)
knitr::opts_chunk$set(error = TRUE,
  upload.fun = imgur_upload, exercise.cap = "Vectors in R")
```

> "*The fundamental data type in R is the vector.*" - THE ART OF R PROGRAMMING, 2011 by Norman Matloff.

## Atomic vectors 

"Atomic" here means *'of or forming a single irreducible unit or component in a larger system.'* Atomic vectors are the basic building blocks of data objects in R. 

There are four types of atomic vectors: logical, integer, double, and character.

### Logical vectors

Logical vectors can only take two values: `TRUE` and `FALSE`, and we can create a logical vector using the `c()` function (which stands for 'combine').

```{r vector_logical, exercise=TRUE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
# print 'vector_logical' below

```

```{r prepare-vector_logical-typeof, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
```

We're going to be using the `class()` and `typeof()` functions to investigate these different objects. 

Check the type and class of `vector_logical` with `typeof()` and `class()`

```{r vector_logical-typeof, exercise=TRUE, exercise.setup = "prepare-vector_logical-typeof"}
# check the type of 'vector_logical' below
typeof()
# check the class of 'vector_logical' below
class()
```

Logical vectors are handy because they add up to the number of `TRUE` values (`TRUE` = 1, `FALSE` = 0).

```{r prepare-vector_logical-sum, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3, 7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

Get the `sum()` of `vector_logical` below:

```{r sum-vector_logical, exercise=TRUE, exercise.setup = "prepare-vector_logical-sum"}
# check the sum of 'vector_logical' below
sum()
```

Logical vectors are also helpful for subsetting, which we will cover in future sections. 

We can also test a vector to see if it's logical with `is.logical()`

```{r prepare-vector_logical-test-is-logical, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3, 7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

```{r vector_logical-test-is-logical, exercise=TRUE, exercise.setup = "prepare-vector_logical-test-is-logical"}
# test 'vector_logical' below
is.logical()
```

Each vector comes with it's own testing function.

## Numeric vectors 

There are two types of numeric vectors, double and integer.

### Integer vectors

Integer vectors are whole numbers, and need to be created with an `L` behind each number. 

```{r integer, exercise=TRUE}
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
# type 'vector_integer' on the line below and click Run Code again

```

```{r prepare-typeof-vector_integer, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3, 7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

Check the type of `vector_integer` with `typeof()` and `class()`

```{r typeof-vector_integer, exercise=TRUE, exercise.setup = "prepare-typeof-vector_integer"}
# check type of vector_integer
typeof()
# check class of vector_integer
class()
```

### Double vectors

Double vectors can have whole numbers *and* decimal points. 

```{r vector_double, exercise=TRUE}
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
# type 'vector_double' on the line below and click Run Code again

```

```{r prepare-vector_double, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

Check the type of `vector_double` with `typeof()`.

```{r typeof-vector_double, exercise=TRUE, exercise.setup = "prepare-vector_double"}
# check type of vector_double
typeof()
```

### Testing numeric vectors

We can test integer and double vectors with `is.numeric()`:

```{r prepare-is-numeric, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

Test the `vector_double` and `vector_integer` with `is.numeric()` below

```{r is-numeric, exercise=TRUE, exercise.setup = "prepare-is-numeric"}
# check vector_integer with is.numeric()
is.numeric()
# check vector_double with is.numeric()
is.numeric()
```

Because these only test if the vector is numeric, we can be more specific with the `is.integer()` function. 

Check `vector_integer` with `is.integer()`

```{r prepare-is-integer-vector_integer, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

```{r is-integer-vector_integer, exercise=TRUE, exercise.setup = "prepare-is-integer-vector_integer"}
# check vector_integer with is.integer()
is.integer()
```

Check `vector_double` with `is.double()`

```{r prepare-is-double-vector_double, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

```{r is-double-vector_double, exercise=TRUE, exercise.setup = "prepare-is-double-vector_double"}
# check vector_double with is.double()
is.double()
```

## String vectors

The final kind of atomic vector is a character (or string) vector, and these contain text information:

```{r vector_string, exercise=TRUE}
# type 'vector_string' on the line below and click Run Code again
vector_string <- c("hat", "chair", "cat", "dog", "book")

```

```{r prepare-typeof-vector_string, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

Check the type of `vector_string` below:

```{r typeof-vector_string, exercise=TRUE, exercise.setup = "prepare-typeof-vector_string"}
# check type of vector_string
typeof()
```

Check `vector_string` with `is.character()`

```{r prepare-is-character-vector_string, include=FALSE}
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_double <- c(3.5, 0.7, 29, 4, 6.78)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_string <- c("hat", "chair", "cat", "dog", "book")
```

```{r is-character-vector_string, exercise=TRUE, exercise.setup = "prepare-is-character-vector_string"}
# check vector_string with is.character()
is.character()
```

## Conclusion

We've covered atomic vectors in these exercises--next up we will review R's S3 atomic vectors.

<div data-iframe-height></div>