---
comment: "*subsetting data in R*"
output: 
  html_document:
    includes:
      in_header: resizer.html
       
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(skimr)
library(knitr)
knitr::opts_chunk$set(error = TRUE,
  upload.fun = imgur_upload, exercise.cap = "extracting data from objects in R")
```

There are three primary methods for subetting in R: 

1. single brackets: `[`  
2. double brackets: `[[`  
3. dollar sign: `$`  

```{r prepare-num_vec, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

View the numeric vector (`num_vec`) in the code chunk by clicking on the **Run Code** button:

```{r num_vec, exercise=TRUE, exercise.setup = "prepare-num_vec"}
num_vec
```

Subset `num_vec` using brackets and a numerical position (or index):

```{r prepare-num_vec-index, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index, exercise=TRUE, exercise.setup = "prepare-num_vec-index"}
# subset the number at position 5

# subset the number at position 2

```

We can also subset a range of numbers using `:`. Subset the numbers from position 2-4 using `2:4`:

```{r prepare-num_vec-index-range, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-range, exercise=TRUE, exercise.setup = "prepare-num_vec-index-range"}
# subset numbers 2-4 

```

The output from `num_vec[2:4]` is another vector (of the numbers we've subsetted).

```{r prepare-num_vec-index-range-str, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-range-str, exercise=TRUE, exercise.setup = "prepare-num_vec-index-range"}
# check the structure of num_vec[2:4] with str() below
str()
```

This is a numeric vector with three numbers (`num [1:3] 9 4 3`)

We can also use the `c()` function to create an index and use it to subset `num_vec`. The object `x` below contains the three positions for the numerical index above (`2:4`).

```{r prepare-num_vec-index-c, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-c, exercise=TRUE, exercise.setup = "prepare-num_vec-index-c"}
# create x with c()
x <- c(2, 3, 4)
# subset num_vec with x

```

We can see the `num_vec[x]` returns the same result as `num_vec[2:4]`. 

We can also use double-brackets to subset a single position from a vectors (`[[]]`), but this doesn't work for a numerical range: 

```{r prepare-num_vec-compare-brackets, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-compare-brackets, exercise=TRUE, exercise.setup = "prepare-num_vec-index-c", error=TRUE}
# this works
str(num_vec[[2]])
# this doesn't
str(num_vec[[2:4]])
```

Double brackets are usually reserved for subsetting more complex objects (like `data.frame`s or `list`s).


```{r prepare-MovieTibble-str, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

Most of the data we'll encounter in this course will be in `data.frame`s or `tibble`s, so we've loaded the `MovieTibble` `tibble`. Check the structure below: 

```{r MovieTibble-str, exercise=TRUE, exercise.setup = "prepare-MovieTibble-str"}
# check the structure of MovieTibble
str()
```

`tibbles` are rectangular (like spreadsheets) and contain vectors of equal length. The dollar sign (`$`) can be used to subset the vectors by name.

Use the dollar sign below to subset the `title` column from `MovieTibble`, but wrap the output in the `head()` function to minimize the output.

```{r prepare-MovieTibble-dollar-sign, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-dollar-sign, exercise=TRUE, exercise.setup = "prepare-MovieTibble-dollar-sign"}
# view the 'head()' of the title column
head()
```

**If we use the dollar sign to subset columns from a `tibble` or `data.frame`, the output is a vector.**

We can also use the row and column position to subset `tibble`s just like vectors. But in order to use brackets, we need to supply **two positions** (or indexes): One for the row, and another for the column (as `data[row, column]`).

Use a single bracket to get the `length` of the movie `13 Going On 30`. 

```{r prepare-MovieTibble-bracket, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-bracket, exercise=TRUE, exercise.setup = "prepare-MovieTibble-bracket"}
# subset the 'length' (column 3) for '13 Going On 30' (in row 5)
MovieTibble[ , ]
```

Note that if we provide two indexes (`[ 5, 3]`) to subset `MovieTibble`, the return object is a `tibble`. The same is true if we use a single row or column index: 

```{r prepare-MovieTibble-bracket-single-row-col-index, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-bracket-single-row-col-index, exercise=TRUE, exercise.setup = "prepare-MovieTibble-bracket-single-row-col-index"}
# single row index
str(MovieTibble[ 5, ])
# single column index
str(MovieTibble[ , 3])
```

However, if we convert `MovieTibble` to a `data.frame` with `as.data.frame()`, a single row index returns a `data.frame` (with all seven columns), but a single column index returns a vector (of all rows): 

```{r prepare-MovieDataFrame-bracket-single-row-index, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieDataFrame-bracket-single-row-index, exercise=TRUE, exercise.setup = "prepare-MovieDataFrame-bracket-single-row-index"}
# convert to data.frame
MovieDataFrame <- as.data.frame(MovieTibble)
# provide a single row index for '13 Going On 30' (row 5)
str(MovieDataFrame[ , ])
# provide a single column index for the 'length' (column 3)
str(MovieDataFrame[ , ])
```

This distinction is important because we need to know the structure of each object we're subsetting. 

We can use column names to subset, too. We can use `c()` to subset the `"title"` and `"genres"` for `13 Going On 30`. 

```{r prepare-MovieTibble-bracket-c-named-cols, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-bracket-c-named-cols, exercise=TRUE, exercise.setup = "prepare-MovieTibble-bracket-c-named-cols"}
# enter "title" and "genres" in the c() to subset 
# these columns
MovieTibble[ 5, c("", "")]
```

We can also use this method to subset the `MovieDataFrame`: 

```{r prepare-MovieDataFrame-bracket-c-named-cols, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieDataFrame-bracket-c-named-cols, exercise=TRUE, exercise.setup = "prepare-MovieDataFrame-bracket-c-named-cols"}
# enter "title" and "genres" in the c() to subset 
# these columns
MovieDataFrame[ 5, c("", "")]
```

Note the difference in the printed output: the `tibble` gives us the dimensions of the output object (`# A tibble: 1 × 2`), and the *number of rows* (`1`), while the `data.frame` prints column names (without the format) the row index (`5`).

Recall that the dollar-sign returns a vector (`MovieTibble$`), which means we can subset this output by combing it with brackets (`[]`)

```{r prepare-MovieTibble-dollar-sign-bracket, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-dollar-sign-bracket, exercise=TRUE, exercise.setup = "prepare-MovieTibble-dollar-sign-bracket"}
# subset the movie "13 Going On 30" (in position 5)
MovieTibble$title[ ]
```

We can also combine the dollar sign `$` with `==` to return a logical vector. Use `str()` to view the structure of the two logical conditions below:

```{r prepare-MovieTibble-dollar-sign-equals, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieTibble-dollar-sign-equals, exercise=TRUE, exercise.setup = "prepare-MovieTibble-dollar-sign-equals"}
# return the structure of movies less than 95 minutes long
str(MovieTibble$length < )
```

```{r prepare-MovieDataFrame-dollar-sign-equals, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieDataFrame-dollar-sign-equals, exercise=TRUE, exercise.setup = "prepare-MovieDataFrame-dollar-sign-equals"}
# return the structure of movies greater than 90 minutes long
str(MovieDataFrame$length > )
```

Logical vectors might not seem very helpful, but they come in handy when we combine them with `[]`.

We can use the logical vectors from `$` and `==` in combination with `[]` to subset `data.frame`s and `tibble`s. Use the code below to subset all movies in `MovieTibble` that are 120 minutes long.

```{r prepare-MovieTibble-dollar-sign-equals-bracket, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
```

```{r MovieTibble-dollar-sign-equals-bracket, exercise=TRUE, exercise.setup = "prepare-MovieTibble-dollar-sign-equals-bracket"}
# return movies that are 120 minutes long (recall that 'length' is 
# a column, so the '120' should be in the column position)
MovieTibble[MovieTibble$length == , ]
```

We can also use `c()` with `$`, `==` and `[]` to subset multiple movie lengths from `MovieDataFrame`.

```{r prepare-MovieDataFrame-dollar-sign-equals-bracket-c, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieDataFrame-dollar-sign-equals-bracket-c, exercise=TRUE, exercise.setup = "prepare-MovieDataFrame-dollar-sign-equals-bracket-c"}
# return movie 'title' and 'length' that are 85 and 110 minutes long 
MovieDataFrame[MovieDataFrame$length == c( , ), c(" ", " ")]
```

We can control the subsetted results for *both* tibbles and data frames using double brackets (`[[]]`). Let's review the behavior of single brackets. If we use a single bracket (without commas) and a numerical index, we get the first column in both `MovieTibble` and `MovieDataFrame` as a rectangular object.

```{r prepare-MovieTibble-MovieDataFrame-review-bracket, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieTibble-MovieDataFrame-review-bracket, exercise=TRUE, exercise.setup = "prepare-MovieTibble-MovieDataFrame-review-bracket"}
# character column from tibble
str(MovieTibble[1])
# character column from data frame 
str(MovieDataFrame[1])
```

But if we use double-brackets, we get the same first column, but as a vector. 

```{r prepare-MovieTibble-MovieDataFrame-double-brackets, include=FALSE}
source("https://bit.ly/3c6DfGe")
MovieTibble <- create_movie_data()
MovieDataFrame <- as.data.frame(MovieTibble)
```

```{r MovieTibble-MovieDataFrame-double-brackets, exercise=TRUE, exercise.setup = "prepare-MovieTibble-MovieDataFrame-double-brackets"}
# character column as vectors
str(MovieTibble[[1]])
str(MovieDataFrame[[1]])
```

<div data-iframe-height></div>