---
comment: "*subsetting data in R*"
output: 
  html_document:
    includes:
      in_header: resizer.html
       
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(skimr)
library(knitr)
knitr::opts_chunk$set(error = TRUE,
  upload.fun = imgur_upload, exercise.cap = "extracting data from objects in R")
```

There are three primary methods for subetting in R: 

1. single brackets: `[`  
2. double brackets: `[[`  
3. dollar sign: `$`  

```{r prepare-num_vec, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

View the numeric vector (`num_vec`) in the code chunk by clicking on the **Run Code** button:

```{r num_vec, exercise=TRUE, exercise.setup = "prepare-num_vec"}
num_vec
```

Subset `num_vec` using brackets and a numerical position (or index):

```{r prepare-num_vec-index, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index, exercise=TRUE, exercise.setup = "prepare-num_vec-index"}
# subset the number at position 5

# subset the number at position 2

```

We can also subset a range of numbers using `:`. Subset the numbers from position 2-4 using `2:4`:

```{r prepare-num_vec-index-range, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-range, exercise=TRUE, exercise.setup = "prepare-num_vec-index-range"}
# subset numbers 2-4 

```

The output from `num_vec[2:4]` is another vector (of the numbers we've subsetted).

```{r prepare-num_vec-index-range-str, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-range-str, exercise=TRUE, exercise.setup = "prepare-num_vec-index-range"}
# check the structure of num_vec[2:4] with str() below
str()
```

This is a numeric vector with three numbers (`num [1:3] 9 4 3`)

We can also use the `c()` function to create an index and use it to subset `num_vec`. The object `x` below contains the three positions for the numerical index above (`2:4`).

```{r prepare-num_vec-index-c, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-index-c, exercise=TRUE, exercise.setup = "prepare-num_vec-index-c"}
# create x with c()
x <- c(2, 3, 4)
# subset num_vec with x

```

We can see the `num_vec[x]` returns the same result as `num_vec[2:4]`. 

We can also use double-brackets to subset vectors (`[[]]`), but this is usually reserved for subsetting more complex objects (like `data.frame`s or `list`s).

```{r prepare-num_vec-compare-brackets, include=FALSE}
num_vec <- c(2, 9, 4, 3, 7)
```

```{r num_vec-compare-brackets, exercise=TRUE, exercise.setup = "prepare-num_vec-index-c"}
# the outputs below are identical
str(num_vec[[2:4]])
str(num_vec[2:4])
```



```{r prepare-MovieData-str, include=FALSE}
source("MovieData.R")
```




