---
title: "*Data import with readr, readxl, and googlesheets4 cheatsheet*"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    df_print: default
    includes:
      in_header: resizer.html
runtime: shiny_prerendered
description: >
  Import data from .csv, .xlsx, and googlesheets.
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(skimr)
library(knitr)
knitr::opts_chunk$set(error = TRUE,
  upload.fun = imgur_upload, 
  exercise.cap = "import cheatseet")
```

> "The tidyverse provides several packages for importing data into R and this cheatsheet covers three of them."


## The cheatsheet 

Access the cheatsheet [here](https://github.com/rstudio/cheatsheets/blob/main/data-import.pdf).

This cheatsheet covers importing data of various formats into RStudio. We're going to show how this looks on your machine (with examples). 

These exercises assume you have a `data/` folder in your current working directory, and that you're (hopefully) using RStudio projects. 

You can download the data uses in this example [here.]() 

## Read Tabular Data with `readr` 

The readr package is part of the `tidyverse`, which we will load below:

```{r packages, eval=FALSE}
library(tidyverse)
```

The data file locations are printed below: 

```{r , eval=FALSE}
# data/
#   ├── file.csv
#   ├── file.tsv
#   ├── file.txt
#   └── file2.csv
```


### Read delimited files 

> " Read files with any delimiter. If no delimiter is specified, it will automatically guess"

This will import the data from the folder path below: 

```{r read_delim-file-txt, exercise=TRUE}
readr::read_delim(file = "data/file.txt")
```



### Useful Read Arguments

## Column Specification with `readr`

## Save Data with `readr`

## Import spreadsheets with `readxl`

## Import spreadsheets with `googlesheets4`

## Conclusion

We've covered atomic vectors in these exercises--next up we will review R's S3 atomic vectors.

<div data-iframe-height></div>