---
comment: "*Rectangular data in R*"
output: 
  html_document:
    includes:
      in_header: resizer.html
       
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(skimr)
library(knitr)
knitr::opts_chunk$set(error = TRUE,
  upload.fun = imgur_upload, exercise.cap = "data.frames and tibbles in R")
```

Most of the time we're encountering data in tables (like spreadsheets). In R, tabular data are stored in `data.frame`s or `tibble`s. 

The code below builds a `data.frame` with the vectors we created in the previous exercise. Click on the **Run Code** button below:

```{r prepare-MyDataFrame, include=FALSE}
vector_string <-  c("hat", "chair", "cat", "dog", "book")
vector_logical <- c(TRUE, FALSE, TRUE, FALSE, FALSE)
vector_integer <- c(4L, 21L, 2L, 3L, 10L)
vector_double <- c(3, 7, 29, 4, 6.78)
```

```{r MyDataFrame, exercise=TRUE, exercise.setup = "prepare-MyDataFrame"}
# this is how we create a data.frame
MyDataFrame <- data.frame(
  "string" = vector_string,
  "logical" = vector_logical,
  "integer" = vector_integer,
  "double" = vector_double
)
```

`data.frame`s in R are similar to spreadsheets. The data are organized in rows and columns. 

Place `MyDataFrame` inside the `str()` function and click on the **Run Code** button below to see how it works. 

```{r prepare-MyDataFrame-str, include=FALSE}
MyDataFrame <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"), 
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = "data.frame", 
    row.names = c(NA, -5L))
```

```{r str-MyDataFrame, exercise=TRUE, exercise.setup = "prepare-MyDataFrame-str"}
# put the MyDataFrame inside the str() function
str()
```

We can see `str()` prints the number of rows (`5 obs.`) and columns (`4 variables:`), and a preview of it's contents. 

In this course, we will primarily use `tibble`s, which are similar to `data.frame`s. The code below creates a `tibble` with the same content as `MyDataFrame`. 

```{r MyTibble, exercise=TRUE}
# this is how we create a tibble
MyTibble <- tibble::tribble(
            ~string, ~logical, ~integer, ~double,
              "hat",     TRUE,       4L,       3,
            "chair",    FALSE,      21L,       7,
              "cat",     TRUE,       2L,      29,
              "dog",    FALSE,       3L,       4,
             "book",    FALSE,      10L,    6.78
            )
```

```{r prepare-MyTibble-str, include=FALSE}
MyTibble <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"),
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = c("tbl_df", "tbl", "data.frame"), 
    row.names = c(NA, -5L))
```

Place `MyTibble` inside the `str()` function and click on the **Run Code** button below to view it's structure:

```{r str-MyTibble, exercise=TRUE, exercise.setup = "prepare-MyTibble-str"}
# put the MyTibble inside the str() function
str()
```

The output from `str(MyTibble)` prints the number of rows x columns (`[5 × 4]`) and the *class* (`S3: tbl_df/tbl/data.frame`), and a preview of it's contents. 

`tibble`s print to the screen a little nicer to the screen. Print `MyTibble` in the code chunk below: 

```{r prepare-MyTibble-print, include=FALSE}
MyTibble <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"),
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = c("tbl_df", "tbl", "data.frame"), 
    row.names = c(NA, -5L))
```

```{r MyTibble-print, exercise=TRUE, exercise.setup = "prepare-MyTibble-print"}
# print MyTibble

```

We can see the format of each column is also printed with the output (`<chr>`, `<lgl>`, `<int>`, `<dbl>`). Compare this to `MyDataFrame` in the code chunk below:

```{r prepare-MyDataFrame-print, include=FALSE}
MyDataFrame <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"), 
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = "data.frame", 
    row.names = c(NA, -5L))
```

```{r MyDataFrame-print, exercise=TRUE, exercise.setup = "prepare-MyDataFrame-print"}
# print MyDataFrame

```

Both `data.frame`s and `tibble`s are lists. Click on the **Run Code** button below to view the `typeof()` for `MyDataFrame` and `MyTibble`:

```{r prepare-MyDataFrame-MyTibble-typeof, include=FALSE}
MyDataFrame <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"), 
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = "data.frame", 
    row.names = c(NA, -5L))
MyTibble <- structure(list(
    string = c("hat", "chair", "cat", "dog", "book"),
    logical = c(TRUE, FALSE, TRUE, FALSE, FALSE), 
    integer = c(4L, 21L, 2L, 3L, 10L), 
    double = c(3, 7, 29, 4, 6.78)), 
    class = c("tbl_df", "tbl", "data.frame"), 
    row.names = c(NA, -5L))
```

```{r MyDataFrame-MyTibble-typeof, exercise=TRUE, exercise.setup = "prepare-MyDataFrame-MyTibble-typeof"}
# place MyDataFrame in typeof()
typeof()
# place MyTibble in typeof()
typeof()
```

